---
title: "CIs and T-tests"
author: "Karla Garibay Garcia"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 1. Confidence intervals


- In 95% of samples that we took from the population we expect to find the population mean within the range for the CI presented. If we took a hundred samples and we found the CI for all of those, then in 95% of the CI presented, we'd actually expect to find the population mean.


- How to describe?The mean length of sea otters (n= 9) is 42.22 inches w/SD of X, 95% CI is 34.7-49.8 inches. 


- While the t.test does get us a meaningful mean and CI, the reported t.test outcome is not meaningful on its own because the alternative hypothesis is the actual output for a 1 sample hypothesis test where the default setting is that a null hypothesis is a mean value equal to 0. Based on our sample and our sample mean, it's very unlikely that we could've drawn our sample from a population of otters that have a mean length equal to 0.


```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

t.test(otter_length)


```

## 2. T-test (1-sample)

In the console, type `?t.test` to have information about the function's default settings in the Help tab. 


See a claim: mean otter length is 50 inches. 

- Test claim using `x = sample` and null hypothesis `mu =`

- To view data in document, call it by typing the name of it in the r code chunk

- What does p-value mean? If the null hypothesis is really 50 inches, there is a 0.0444 chance that I could've randomly taken a sample from that population that has a mean that is at least as different from 50 as my sample mean is in either direction. 


```{r}
otter_test <- t.test(x = otter_length, 
                     mu = 50)
otter_test

```

## 3. T-test (2-sample)


Are my differences in my two sample means so different that I think they were drawn from populations with different means. If they were drawn from populations with the same mean then it's:
- More likely that I will get sample means that are close together taking into account data spread and sample size
- Less likely that I will get sample means that are very far apart 


Welch's 2-sample t-test does not assume equal variances, 


Gives me: 
- Mean of both samples

- What does CI mean? The expected CI for the difference in mean sizes in the horn length. Can be negative because x-y or y-x could be negative.We'd expect the difference between the sample means to be within the CI range for 95% of the samples I took. CI on the difference in means is not usually reported, however. 


Focus of this t-test: Testing a null hypothesis that the difference in means equals 0. 

- P-value tells me: if these samples are drawn from populations with the same mean horn length, which is the null hypothesis, then there's almost a 90% chance that I could've taken 2 random samples from those populations and found samples at least as different as the samples that I have. It's a really good chance that I could've retrieved my samples from populations with the same mean and gotten means from my samples that are at least as different as mine are. I don't think I have enough evidence to reject that null hypothesis that they're drawn from populations with the same means. 

Things to remember: 

- it is the probability, if the null hypothesis is true (for a 2-sample test it'll be that the 2 populations have the same mean), that we could've gotten samples with means that are at least as different than ours are. If that probability is really high, there's a good chance they were drawn  from populations with the same mean. As that probability gets lower, it's less likely that we could've randomly drawn those samples with the same mean. Which means that when that probability gets really low, we'd instead decide that they were drawn from populations with different means.  


```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 40)
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x = desert_bighorns, 
       y = sierra_bighorns)


```

## Part 2: Two-sample t-test

### Basic wrangling

Create a subset that:

- Retains observations for suvs and compact cars

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))
```

### Exploratory data visualization

1. Exploratory histograms to understand the distribution of our observations before we decide the type of analysis or test that's appropriate. Histograms only plot a single variable. It then does the counts that show up on the x-axis for you. 

`facet_wrap(~class)` gives you data for each class within your data. Number of bins can be adjusted so that it reflects the number of observations in a way that is meaningful and useful for us. 

Samples look somewhat bell-shaped in the histograms. 

```{r}
ggplot(data = compact_suv,
       aes(x = cty)) +
  geom_histogram(bins = 15) +
  facet_wrap(~class)
```

2. QQ Plot: A great way to compare distributions to a normal distribution to see how close to normally distributed they are. Plots my actual values for my observations vs. theoretical values if they're perfectly drawn from a normal distribution. So the closer the relationship gets to linear, the closer my sample is to normal distribution.


```{r}
ggplot(data = compact_suv, 
       aes(sample = cty)) + 
  geom_qq() + 
  facet_wrap(~class)
```

### Descriptive statistics

Use `n = n()` to find out sample size because the larger n gets, the closer our sampling distribution of means will get to normally distributed. So even if our population distribution wasn't normally distributed, because of the CLT and the n, it'll be ok because we still know the distribution of means will be normal, so a t-test would still be appropriate to compare these means. 

This is the baseline exploration to do before considering to do a t-test. Along with critically thinking about whether our sample is representative of our population.  

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarise(mean_city = mean(cty), 
            sd_city = sd(cty),
            n = n())
  
car_stats
```

### 2-sample t-test

#### Create vectors containing observations for each sample (city gas mileage)

Name new sample, starting from original dataframe, filter to only include compact cars, and then use pull funciton to only pull the values in the city column and store them as a vector to run a t-test. If I used select function, it'll store it as a dataframe with a single column, which is not what t.test function wants. Do the same for suv's to get second vector. 


```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

t.test(x = compact_sample, suv_sample)
```
Testing a null hypothesis that the means are equal aka samples are drawn from populations with the same mean. I'm testing for what the probability is of getting samples that are at least this different if that null hypothesis is true. This is what p-value tells me. What p-value tells me here is there's close to a 0 chance that I could've drawn samples with means that are at least this different if they were actually taken from populations with the same means. Aka if suvs and compact cars actually had the same city gas mileage, there's close to a 0% chance that I could've randomly taken samples from that population that ended up being at least this different. So it's probably because we drew these samples from populations with different means. 







